---
title: "2014-2022 college crime data"
output: html_notebook
---

This is an analysis of 2014-2021 Clery Act data, plus 2022 data published by individual institutions. This data includes 21 Maryland institutions that have a combined enrollment of at least 1,000 undergraduates and graduates.

The CSV file used was joined and cleaned by The Baltimore Sun, but is not included in this repository. You can download the CSV here. Enrollment data was xxxxx

```{r}
library(tidyverse)
library(janitor)
```

### Load in two data files, clean up a bit
```{r}
enrollment <- read_csv("enrollment_ten_years_v2.csv") %>% clean_names() %>% pivot_longer(cols=c("x2012_13_fte_12_month_enrollment":"x2021_22_fte_12_month_enrollment"), names_to = "year", values_to = "enrollment") %>% mutate(year1 = substr(year, 2,3)) %>% mutate(year2 = substr(year, 7,8))

enrollment <- enrollment %>% mutate(year = paste(year1,year2, sep="")) %>% select(unitid, institution_name, year, enrollment) %>% filter(year != "2013")

enrollment_md <- enrollment %>% group_by(year) %>% summarise(enrollment = sum(enrollment))
```

```{r}
df <- read_csv("college_crime.csv")

df <- df %>% mutate(survey_year = as.character(survey_year))

df <- df %>% full_join(enrollment, by=c("survey_year"="year","institution_me"="institution_name"))
```
#### Integrity checks
## How many survey years?
```{r}
df %>% group_by(survey_year) %>% summarise(count = n())
```
2014-2022

## How many institution names?
```{r}
df %>% group_by(institution_me) %>% summarise(count = n()) %>% arrange(desc(count))
```
21 schools at several different counts for how many

## Looking at Johns Hopkins for an example of how campuses and institution sizes are organized
```{r}
df %>% filter(institution_me == "Johns Hopkins University") %>% select(survey_year, institution_me, campus_me, enrollment) %>% group_by(survey_year, institution_me, campus_me, enrollment) %>% summarise(count = n())
```
## By category
```{r}
df %>% group_by(category) %>% summarise(count = n())
```
3 categories
172 non campus
282 on campus
282 public property

## Campus name
```{r}
df %>% group_by(institution_me) %>% summarise(count = n()) %>% arrange(desc(count))

df %>% group_by(institution_me, campus_me) %>% summarise(count = n()) %>% arrange(desc(count))
```

32 campus names across 21 campuses

## JHU example
```{r}
jhu <- df %>% filter(institution_me == "Johns Hopkins University")

jhu %>% group_by(campus_me) %>% summarise(count = n())
```
Six campuses

#### Analysis
## Creating subset of MD 2014-2022 total, violent, sexual, property crime from main df 
```{r}
total_crime_md <- 
  df %>% 
  group_by(survey_year) %>% 
  summarise(total_crime = sum(total_crime))

violent_crime_md <- 
  df %>% 
  group_by(survey_year) %>% 
  summarise(violent_crime = sum(violent_crime))

sexual_crime_md <- 
  df %>% 
  group_by(survey_year) %>% 
  summarise(sexual_crime = sum(sexual_crime))

property_crime_md <- 
  df %>% mutate(property_crime = arson+motor_vehicle_theft+burglary) 
  
property_crime_md <- 
  df %>% 
  group_by(survey_year) %>% 
  summarise(property_crime = sum(property_crime))
```

## Join with enrollment table
```{r}
total_crime_md <- total_crime_md %>% inner_join(enrollment_md, by=c("survey_year"="year"))

violent_crime_md <- violent_crime_md %>% inner_join(enrollment_md, by=c("survey_year"="year"))

sexual_crime_md <- sexual_crime_md %>% inner_join(enrollment_md, by=c("survey_year"="year"))

propety_crime_md <- property_crime_md %>% inner_join(enrollment_md, by=c("survey_year"="year"))
```

## Calculate rates
```{r}
total_crime_md <- total_crime_md %>% mutate(total_crime_rate = total_crime/enrollment*1000)

violent_crime_md <- violent_crime_md %>% mutate(violent_crime_rate = violent_crime/enrollment*1000)

sexual_crime_md <- sexual_crime_md %>% mutate(sexual_crime_rate = sexual_crime/enrollment*1000)

property_crime_md <- propety_crime_md %>% mutate(property_crime_rate = property_crime/enrollment*1000)
```

### 2022 average rates by university
## Filtering by 2022 for each crime category and adding for each university
```{r}
df <- df %>% mutate(survey_year = as.numeric(survey_year)) 

total_crime_uni_2022 <- df %>% filter(survey_year == "2022") %>% group_by(survey_year, institution_me) %>% summarise(total_crime = sum(total_crime)) %>% mutate(survey_year = as.character(survey_year))

violent_crime_uni_2022 <- df %>% filter(survey_year == "2022") %>% group_by(survey_year, institution_me) %>% summarise(violent_crime = sum(violent_crime)) %>% mutate(survey_year = as.character(survey_year))

sexual_crime_uni_2022 <- df %>% filter(survey_year == "2022") %>% group_by(survey_year, institution_me) %>% summarise(sexual_crime = sum(sexual_crime)) %>% mutate(survey_year = as.character(survey_year))
```
## Join with enrollment and calculate rate
```{r}
total_crime_uni_2022 <- total_crime_uni_2022 %>% inner_join(enrollment, by=c("survey_year"="year","institution_me"="institution_name"))

total_crime_uni_2022 <- total_crime_uni_2022 %>% mutate(total_crime_rate = total_crime/enrollment*1000)


violent_crime_uni_2022 <- violent_crime_uni_2022 %>% inner_join(enrollment, by=c("survey_year"="year","institution_me"="institution_name"))

violent_crime_uni_2022 <- violent_crime_uni_2022 %>% mutate(violent_crime_rate = violent_crime/enrollment*1000)


sexual_crime_uni_2022  <- sexual_crime_uni_2022  %>% inner_join(enrollment, by=c("survey_year"="year","institution_me"="institution_name"))

sexual_crime_uni_2022  <- sexual_crime_uni_2022  %>% mutate(sexual_crime_rate = sexual_crime/enrollment*1000)
```

### Average rates by university, 2014-2022
## Group by year and college for each crime category
```{r}
total_crime_uni <- df %>% group_by(survey_year, institution_me) %>% summarise(total_crime = sum(total_crime)) 

violent_crime_uni <- df %>% group_by(survey_year, institution_me) %>% summarise(violent_crime = sum(violent_crime)) 

sexual_crime_uni <- df %>% group_by(survey_year, institution_me) %>% summarise(sexual_crime = sum(sexual_crime))
```

## Join enrollment, calculate rates
```{r}
total_crime_uni <- total_crime_uni %>% mutate(survey_year = as.character(survey_year)) %>% inner_join(enrollment,by=c("survey_year"="year","institution_me"="institution_name"))

total_crime_uni <- total_crime_uni %>% mutate(total_crime_rate = total_crime/enrollment*1000)


violent_crime_uni <- violent_crime_uni %>% mutate(survey_year = as.character(survey_year))%>% inner_join(enrollment, by=c("survey_year"="year","institution_me"="institution_name"))

violent_crime_uni <- violent_crime_uni %>% mutate(violent_crime_rate = violent_crime/enrollment*1000)


sexual_crime_uni  <- sexual_crime_uni  %>% mutate(survey_year = as.character(survey_year)) %>% inner_join(enrollment,by=c("survey_year"="year","institution_me"="institution_name"))

sexual_crime_uni  <- sexual_crime_uni  %>% mutate(sexual_crime_rate = sexual_crime/enrollment*1000)
```

## 
```{r}
sum_by_type_md <- df %>% group_by(survey_year) %>% 
  summarise(sum_violent_crime = sum(violent_crime),
            sum_murder = sum(murder_non_negligent_manslaughter),
            sum_neg_manslaughter = sum(negligent_manslaughter),
            sum_rape = sum(rape),
            sum_fondling = sum(fondling),
            sum_incest = sum(incest),
            sum_statutory = sum(statutory_rape),
            sum_robbery = sum(robbery), 
            sum_agg_assault = sum(aggravated_assault),
            sum_burglary = sum(burglary),
            sum_motor_vehicle = sum(motor_vehicle_theft),
            sum_arson = sum(arson)) 


sum_by_type_md <- sum_by_type_md %>% inner_join(enrollment_md, by=c("survey_year"="year"))


sum_by_type_md <- sum_by_type_md %>% mutate(violent_crime_rate = sum_violent_crime/enrollment*1000) %>%
  mutate(murder_rate = sum_murder/enrollment*1000) %>%
  mutate(neg_manslaughter_rate = sum_neg_manslaughter/enrollment*1000) %>%
  mutate(fondling_rate = sum_fondling/enrollment*1000) %>%
  mutate(incest_rate = sum_incest/enrollment*1000) %>%
  mutate(statutory_rate = sum_statutory/enrollment*1000) %>%
  mutate(murder_rate = sum_murder/enrollment*1000) %>%
  mutate(robbery_rate = sum_robbery/enrollment*1000) %>%
  mutate(agg_rate = sum_agg_assault/enrollment*1000) %>%
  mutate(rape_rate = sum_rape/enrollment*1000) %>% 
  mutate(burglary_rate = sum_burglary/enrollment*1000) %>%
  mutate(motor_vehicle_rate = sum_motor_vehicle/enrollment*1000) %>%
  mutate(arson_rate = sum_arson/enrollment*1000)
```
## Percentage by violent crime type, Maryland 2022
```{r}
df %>% group_by(survey_year) %>% summarise(sum_violent_crime = sum(violent_crime), sum_murder = sum(murder_non_negligent_manslaughter), sum_rape = sum(rape), sum_robbery = sum(robbery), sum_agg_assault = sum(aggravated_assault)) %>% mutate(pct_murder = sum_murder/sum_violent_crime*100, pct_rape = sum_rape/sum_violent_crime*100, pct_robbery = sum_robbery/sum_violent_crime*100, pct_agg_assault = sum_agg_assault/sum_violent_crime*100)
```
## Calculating count for rape without the NAs - added 8 to make 108 in 2022 to account for the 8 forcible sex offenses at Bowie and UMD Eastern Shore
```{r}
df %>% filter(!is.na(rape)) %>% filter(survey_year == "2022") %>% 
  group_by(survey_year) %>% 
  summarise(sum_rape = sum(rape))
```
## Percent of each crime by college
```{r}
sums_2022_by_uni <- df %>% filter(survey_year == "2022") %>% group_by(institution_me) %>% summarise(sum_total_crime = sum(total_crime), sum_violent_crime = sum(violent_crime), sum_property_crime = sum(property_crime), sum_sexual_crime = sum(sexual_crime))

pct_2022_by_uni <- sums_2022_by_uni %>% mutate(pct_violent = sum_violent_crime/sum_total_crime*100) %>% mutate(pct_property = sum_property_crime/sum_total_crime*100) %>% mutate(pct_sexual = sum_sexual_crime/sum_total_crime*100) 
```

### Aggravated assault - Share of aggravated assasults has grown to nearly 50% in 2022

## Create table of aggravated assault rates by year and by college
```{r}
agg_assault_crime_uni <- df %>% group_by(survey_year, institution_me) %>% summarise(ag_crime = sum(aggravated_assault)) %>% mutate(survey_year = as.character(survey_year))

agg_assault_crime_uni  <- agg_assault_crime_uni %>% inner_join(enrollment, by=c("survey_year"="year","institution_me"="institution_name"))

agg_assault_crime_uni  <- agg_assault_crime_uni %>% mutate(ag_crime_rate = ag_crime/enrollment*1000)

agg_assault_crime_uni <- agg_assault_crime_uni %>% select(survey_year, institution_me, ag_crime_rate) %>% pivot_wider(names_from = survey_year, values_from = ag_crime_rate)
```

## Who has the most aggravated assaults, raw numbers
```{r}
agg_assault_2022 <- df %>% filter(survey_year == "2022") %>% group_by(institution_me) %>% summarise(sum = sum(aggravated_assault))
```
## Looking at JHU raw aggravated assaults
```{r}
agg_assault_jhu <- df %>% filter(institution_me == "Johns Hopkins University") %>% group_by(survey_year, institution_me) %>% summarise(jhu_agg_assaults = sum(aggravated_assault))
```

## Aggravated assaults by on campus, non campus, public property
```{r}
agg_assault <- df %>% group_by(survey_year, category) %>% summarise(sum_agg = sum(aggravated_assault))

agg_assault <- agg_assault %>% pivot_wider(names_from = category, values_from = sum_agg)
```

```{r}
agg_assault <- agg_assault %>% mutate(agg = sum(`non-campus`+`public property`+`on campus`)) %>% mutate(pct_on_campus = `on campus`/`agg`*100)
```

## Bowie aggravated assaults
```{r}
agg_assault_bowie <- df %>% filter(institution_me == "Bowie State University") %>% group_by(survey_year, institution_me) %>% summarise(ag_bowie= sum(aggravated_assault))
```
## UMD Ccollege Park aggravated assaults
```{r}
ag_assault_cp <- df %>% filter(institution_me == "University of Maryland-College Park") %>% group_by(survey_year, institution_me) %>% summarise(ag_cp = sum(aggravated_assault))
```
## Raw aggravated assaults by college and by year
```{r}
ag_assault_uni_raw <- df %>% group_by(survey_year, institution_me) %>% summarise(ag_assault = sum(aggravated_assault)) %>% pivot_wider(names_from = survey_year, values_from = ag_assault)
```

## JHU aggravated assault breakdown by campus
```{r}
jhu_agg_campuses <- df %>% filter(institution_me == "Johns Hopkins University") %>% filter(category == "on campus") %>% group_by(survey_year, campus_me) %>% summarise(sum = sum(aggravated_assault))

jhu_agg_campuses <- jhu_agg_campuses %>% pivot_wider(names_from = survey_year, values_from = sum)
```

## Aggravated assaults at JHU by category
```{r}
df %>% filter(survey_year == "2022") %>% filter(institution_me == "Johns Hopkins University") %>% group_by(category) %>% summarise(aggravated_assault = sum(aggravated_assault))

```
## Aggravated assaults at JHU by campus, on campus only
```{r}
df %>% filter(survey_year == "2022") %>% filter(institution_me == "Johns Hopkins University")  %>% filter(category=="on campus") %>% group_by(campus_me) %>% summarise(aggravated_assault = sum(aggravated_assault))
```

## Creating table for Datawrapper chart 1
```{r}
crime_uni_chart <- total_crime_md %>% inner_join(violent_crime_md, by= "survey_year") %>% inner_join(sexual_crime_md, by= "survey_year") %>% inner_join(property_crime_md, by= "survey_year")
```

## Creating table for Datawrapper chart 2
```{r}
violent_crime_chart <- sum_by_type_md %>% select(survey_year, violent_crime_rate, agg_rate, rape_rate, robbery_rate, murder_rate)
```
Rate for rape externally calculated due to 
